---
title: Sending DECIDE Monthly Newsletters
output: html_document
---


This R script is a simple way of sending emails out in bulk, it gets the list of users from google sheets, generates email content for all users (step 1). Then it goes through all the users again and sends the emails (step 2).

This is set up as a markdown document so that it could in theory be run on a schedule on Rstudio: https://docs.rstudio.com/connect/user/scheduling/ connect.

Initial set up, packages, credentials for email / google sheets:

```{r}

library(googlesheets4)
library(blastula)
library(rmarkdown)

creds <- creds_envvar(user = "simonrolph.ukceh@gmail.com",
                      pass_envvar = "gmail_password",
                      provider = "gmail",
                      use_ssl = T)

#authenticate google sheets
gs4_auth(
    cache = ".secrets",
    email = "simonrolph.ukceh@gmail.com"
)

```

Load the users from googlesheets:

```{r}
#load the users
#read as all character with the col_types argument (especially the irecord warehouse ID which it likes to convert to a scientific notation number)
user_db <- range_read("1akEZzgb5tnMNQhnAhH3OftLm0e1kyH8alhCYIHcYxes",col_types = c("c"))

#create a new column which will be used to store the file location of outputted email
user_db$newsletter_file_location <- ""
```

Step 1: generate the newsletters

For each user:
 * create newsletter
 * randomise items
 * save outputs


```{r}
n_users <- nrow(user_db)

#to be looped
#i <- 1

for (i in 1:n_users){
  #get all the markdown parameters into a list
  markdown_params <- 
    list(
        name = user_db$name[i],
        email = user_db$email[i],
        irecord_username = user_db$irecord_username[i],
        ispot_username = user_db$ispot_username[i],
        inat_username = user_db$inat_username[i],
        irecord_key      = gsub("Â","",Sys.getenv("irecord_key")),
        ispot_key        = Sys.getenv("ispot_key")
    )
  
  #if any usernames are not there then set them to NA
  if (length(markdown_params$irecord_username)==0){markdown_params$irecord_username <- NA}
  if (length(markdown_params$ispot_username)==0){markdown_params$ispot_username <- NA}
  if (length(markdown_params$inat_username)==0){markdown_params$inat_username <- NA}
  
  
  print("Generating newsletter")
  
  #create a folder for the day's newsletters
  if(!dir.exists(paste0("newsletters/",Sys.Date()))){
    dir.create(paste0("newsletters/",Sys.Date()))
  }
  
  #render the newsletter
  out_file_name <- paste0("../newsletters/",Sys.Date(),"/",markdown_params$name,".html")
  
  user_db$newsletter_file_location[i] <- render("newsletter_templates/v0_0_1.Rmd",
                output_file = out_file_name,
                params = markdown_params
  )
}





```

Step 2: send the pre-generated templates out to everyone.

for each user:
Send emails

```{r}
#i <- 1

for (i in 1:n_users){
  #define sender and recipient (could change this for testing, eg. set the recipient to your email to generate a batch of emails for different users but then see what they look like in your own inbox without spamming real users.)
  sender <- "simonrolph.ukceh@gmail.com" # obviously we want to change then in real use
  recipients <- user_db$email[i]
  
  # turn the rendered markdown into a blastula ready email object
  email_obj <- blastula:::cid_images(user_db$newsletter_file_location[i])
  
  #send email
  smtp_send(email_obj,
            from = sender,
            to = recipients,
            subject = "Your DECIDE Newsletter",
            credentials = creds,
            verbose = T
  )

}

```








