---
title: Sending DECIDE Monthly Newsletters
output: html_document
---


This R script is a simple way of sending emails out in bulk, it gets the list of users from a csv, builds a dataframe (step 0), generates email content for all users (step 1). Then it goes through all the users again and sends the emails (step 2).

This is set up as a markdown document so that it could in theory be run on a schedule on Rstudio: https://docs.rstudio.com/connect/user/scheduling/ connect.

Initial set up, packages, credentials for email / google sheets:

```{r}
library(blastula)
library(rmarkdown)
library(readr)
library(fst)
library(dplyr)

library(sf)


```

Load the users:

```{r}
#load dataframe
user_db <- read_csv("data_personal/data-2022-06-06.csv")
names(user_db)[1] <- "row"


#create a new column which will be used to store the file location of outputted email
user_db$newsletter_file_location <- ""
```

Step 0: build the dataframe


```{r}
#note all the extra back slashes for escaping
file_location <- "\\\\nerclactdb.adceh.ceh.ac.uk\\appdev\\appdev\\DECIDE\\data\\species_data\\data_cache\\butterfly"

records <- list.files(file_location) %>%
  paste(file_location,.,sep="\\") %>%
  lapply(read_fst) %>%
  bind_rows()

glimpse(records)

#get easting northings
records <- st_as_sf(records,coords = c('longitude', 'latitude'), crs = 4326)  #sf version
records[,c("longitude","latitude")] <- st_coordinates(records)

records  <- records %>% st_transform(27700)
records[,c("easting","northing")] <- st_coordinates(records)
records <- st_drop_geometry(records)

#get day/month/year columns
records$day <- records$observed_on %>% as.Date() %>% format(format = "%d")
records$month <- records$observed_on %>% as.Date() %>% format(format = "%m")
records$year <- records$observed_on %>% as.Date() %>% format(format = "%Y")

saveRDS(records,"data/species_records.RDS")

```


Step 1: generate the newsletters

For each user:
 * create newsletter
 * save outputs
 
 
```{r}
n_users <- nrow(user_db)

#to be looped
#i <- 1

start_date <- Sys.Date()
end_date <- Sys.Date()-30

for (i in 1:n_users){
  
  
  
  #get all the markdown parameters into a list
  markdown_params <- 
    list(
        name = user_db$name[i],
        irecord_username = user_db$irecord_username[i],
        ispot_username = user_db$ispot_username[i],
        inat_username = user_db$inat_username[i],
        start_date = start_date,
        end_date =  end_date,
        try_personalised = TRUE,
        records_data_location = "data/species_records.RDS",
        home_lat = user_db$home_lat[i],
        home_lon = user_db$home_lon[i],
        irecord_key      = gsub("Â","",Sys.getenv("irecord_key")),
        ispot_key        = Sys.getenv("ispot_key")
    )
  
  #if any usernames are not there then set them to NA
  if (length(markdown_params$irecord_username)==0){markdown_params$irecord_username <- NA}
  if (length(markdown_params$ispot_username)==0){markdown_params$ispot_username <- NA}
  if (length(markdown_params$inat_username)==0){markdown_params$inat_username <- NA}
  
  
  print("Generating newsletter")
  
  #create a folder for the day's newsletters
  if(!dir.exists(paste0("newsletters/",Sys.Date()))){
    dir.create(paste0("newsletters/",Sys.Date()))
  }
  
  #render the newsletter
  out_file_name <- paste0("../newsletters/",Sys.Date(),"/",markdown_params$name,".html")
  
  user_db$newsletter_file_location[i] <- render("newsletter_templates/v0_0_2.Rmd",
                output_file = out_file_name,
                params = markdown_params
  )
}


```

Step 2: send the pre-generated templates out to everyone.

for each user:
Send emails

```{r}
#i <- 1

# creds <- creds_envvar(user = "simrol@ceh.ac.uk",
#                           pass_envvar = "outlook_password",
#                           provider = "office365",
#                           use_ssl = T)
# 
# for (i in 1:n_users){
#   #define sender and recipient (could change this for testing, eg. set the recipient to your email to generate a batch of emails for different users but then see what they look like in your own inbox without spamming real users.)
#   sender <- "simrol@ceh.ac.yk" # obviously we want to change then in real use
#   recipients <- user_db$email[i]
#   
#   # turn the rendered markdown into a blastula ready email object
#   email_obj <- blastula:::cid_images(user_db$newsletter_file_location[i])
#   
#   #send email
#   smtp_send(email_obj,
#             from = sender,
#             to = recipients,
#             subject = "Your DECIDE Newsletter",
#             credentials = creds,
#             verbose = T
#   )
# 
# }

```








