---
title: Sending DECIDE Monthly Newsletters
output: html_document
---


This R script is a simple way of sending emails out in bulk, it gets the list of users from a csv, builds a dataframe (step 0), generates email content for all users (step 1). Then it goes through all the users again and sends the emails (step 2).

This is set up as a markdown document so that it could in theory be run on a schedule on Rstudio: https://docs.rstudio.com/connect/user/scheduling/ connect.

Initial set up, packages, credentials for email / google sheets:

```{r}
library(blastula)
library(rmarkdown)
library(readr)
library(fst)
library(dplyr)
library(sf)

```

Load the users:

```{r}
#load dataframe
user_db <- read_csv("data_personal/data-2022-06-23.csv")
names(user_db)[1] <- "user_id"

#for testing with g pigs
#--
g_pigs <- read.csv("data_personal/g_pigs.txt",header = F)[,1]

user_db <- user_db %>% filter(email %in% g_pigs)
#--


#create a new column which will be used to store the file location of outputted email
user_db$newsletter_file_location <- ""

#for TESTING
#all datastories
all_data_stories <- expand.grid(personalised = c(T,F),ds =1:5)[1:9,]

expand.grid.df <- function(...) Reduce(function(...) merge(..., by=NULL), list(...))
emails <- expand.grid.df(all_data_stories,user_db)
```

Step 0: build the records data dataframe


```{r}
#note all the extra back slashes for escaping
file_location <- "\\\\nerclactdb.adceh.ceh.ac.uk\\appdev\\appdev\\DECIDE\\data\\species_data\\data_cache\\butterfly"

records <- list.files(file_location) %>%
  paste(file_location,.,sep="\\") %>%
  lapply(read_fst) %>%
  bind_rows()

glimpse(records)

#get easting northings
records <- st_as_sf(records,coords = c('longitude', 'latitude'), crs = 4326)  #sf version
records[,c("longitude","latitude")] <- st_coordinates(records)

records  <- records %>% st_transform(27700)
records[,c("easting","northing")] <- st_coordinates(records)
records <- st_drop_geometry(records)

#get day/month/year columns
records$day <- records$observed_on %>% as.Date() %>% format(format = "%d")
records$month <- records$observed_on %>% as.Date() %>% format(format = "%m")
records$year <- records$observed_on %>% as.Date() %>% format(format = "%Y")

saveRDS(records,"data/species_records.RDS")

```


Step 1: generate the newsletters

For each user:
 * create newsletter
 * save outputs
 
 
```{r}
n_emails <- nrow(emails)
#n_emails <- 1 
#to be looped
#i <- 1

start_date <- Sys.Date()-30
end_date <- Sys.Date()

for (i in 1:n_emails){
  #get all the markdown parameters into a list
  markdown_params <- 
    list(
        name = emails$name[i],
        irecord_username = emails$irecord_username[i],
        ispot_username = emails$ispot_username[i],
        inat_username = emails$inat_username[i],
        start_date = start_date,
        end_date =  end_date,
        try_personalised = emails$personalised[i],
        records_data_location = "data/species_records.RDS",
        home_lat = emails$home_lat[i],
        home_lon = emails$home_lon[i],
        irecord_key      = gsub("Â","",Sys.getenv("irecord_key")),
        ispot_key        = Sys.getenv("ispot_key"),
        data_story = as.numeric(emails$ds[i])
    )
  
  #if any usernames are not there then set them to NA
  if (length(markdown_params$irecord_username)==0){markdown_params$irecord_username <- NA}
  if (length(markdown_params$ispot_username)==0){markdown_params$ispot_username <- NA}
  if (length(markdown_params$inat_username)==0){markdown_params$inat_username <- NA}
  
  
  print("Generating newsletter")
  
  #create a folder for the day's newsletters
  if(!dir.exists(paste0("newsletters/",Sys.Date()))){
    dir.create(paste0("newsletters/",Sys.Date()))
  }
  
  #render the newsletter
  out_file_name <- paste0("../newsletters/",
                          Sys.Date(),
                          "/",
                          emails$user_id[i],
                          "_",
                          markdown_params$try_personalised,
                          "_",
                          markdown_params$data_story,
                          ".html")
  
  emails$newsletter_file_location[i] <- render(
    "newsletter_templates/v0_0_4.Rmd",
    output_file = out_file_name,
    params = markdown_params,
    output_options = list(self_contained=TRUE,output = "blastula::blastula_email")
  )
}


```

Step 2: send the pre-generated templates out to everyone.

for each user:
Send emails

```{r}
creds <- creds_envvar(user = "simrol@ceh.ac.uk",
                          pass_envvar = "outlook_password",
                          provider = "office365",
                          use_ssl = T)

for (i in 1:n_emails){
  print(i)
  
  #define sender and recipient (could change this for testing, eg. set the recipient to your email to generate a batch of emails for different users but then see what they look like in your own inbox without spamming real users.)
  sender <- "simrol@ceh.ac.uk" # obviously we want to change then in real use

  # turn the rendered markdown into a blastula ready email object
  email_obj <- blastula:::cid_images(emails$newsletter_file_location[i])

  #send email
  smtp_send(email_obj,
            from = sender,
            to = c(emails$email[i]),
            subject = "MyDECIDE",
            credentials = creds,
            verbose = F
  )
  
  Sys.sleep(1)

}

```








